---
title: "Reading in and cleaning the data"
subtitle: "Solutions"
author: "Tom Paskhalis"
date: "29 November, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE}
# Load libraries
library("pdftools")
library("stringr")
```

```{r}
# Read-in the file
draft <- pdftools::pdf_text("../data/draft_withdrawal_agreement_0.pdf")
```

## Challenge 1

**Easy mode** Now extract all the Directives mentioned in the Withdrawal agreement. You can
either use `str_extract_all` function from `stringr` package. Otherwise,
use `gregexpr` function from base R and pass its output to `regmatches`.
The solution should be able to detect such directives as **Directive 92/84/EEC**,
**Directive 2011/64/EU** and **Directive 2008/118/EC**.

**Medium**: Compute a frequency table of directives mentioned and return
the first 10 and their corresponding frequencies.

**Advanced**: Produce a bar chart showing the frequencies of the
top 10 directives.

**Subject expert**: Explore the topics to which the most mentioned directives refer.

```{r answers1}
pattern <- "Directive [0-9]{2,4}/[0-9]{2,3}/[A-Z]{2,3}"

# `base` R solution
directives1 <- regmatches(draft, gregexpr(pattern, draft))

# `stringr` solution
directives2 <- stringr::str_extract_all(draft, pattern)

directives <- unlist(directives1)
directives <- sort(table(directives), decreasing = TRUE)
sum(directives)
directives10 <- directives[1:10]
```

```{r plot1}
barplot(directives10, las = 2)
```

The most mentioned directive, *2004/38/EC* is also called *Citizens’ Rights Directive* or
*Free Movement Directive*. As its name suggests, it covers the rights of the EU
citizens with respect to the freedom of movement. Interestingly, the second most
cited directive is *2001/83/EC* is related to medicinal products for human use.
While the third, *Directive 2005/36/EC* is about the recognition of
professional qualifications.
