---
title: "Keywords-in-Context and Dictionaries"
subtitle: "Solutions"
author: "Tom Paskhalis"
date: "29 November, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE}
# Load libraries
library("readr")
library("dplyr")
library("stringr")
library("lubridate")
library("quanteda")
```

```{r prepare, echo=FALSE, message=FALSE}
# House of Representatives
us_house_2017 <- readr::read_csv("../data/us-house-2017.csv.gz")
us_house_2018 <- readr::read_csv("../data/us-house-2018.csv.gz")
# Senate
us_senate_2017 <- readr::read_csv("../data/us-senate-2017.csv.gz")
us_senate_2018 <- readr::read_csv("../data/us-senate-2018.csv.gz")

pattern <- "20\\d\\d-(0[1-9]|1[012])-(0[1-9]|[12][0-9]|3[01])"
us_house_2017 <- us_house_2017 %>%
    dplyr::mutate(date = lubridate::ymd(stringr::str_extract(granuleId, pattern)))
us_house_2018 <- us_house_2018 %>%
    dplyr::mutate(date = lubridate::ymd(stringr::str_extract(granuleId, pattern)))

congress115 <- dplyr::select(us_house_2017, date, chamber, speaker, text) %>%
  dplyr::bind_rows(dplyr::select(us_house_2018, date, chamber, speaker, text)) %>%
  dplyr::bind_rows(dplyr::select(us_senate_2017, date, chamber, speaker, text)) %>%
  dplyr::bind_rows(dplyr::select(us_senate_2018, date, chamber, speaker, text))

corpus115 <- quanteda::corpus(congress115)

aggregate115 <- congress115 %>%
  dplyr::mutate(year = as.character(lubridate::year(date)),
                month = as.character(lubridate::month(date)),
                year_month_chamber = paste(year, month, chamber, sep = "-")) %>%
  dplyr::group_by(year_month_chamber) %>%
  dplyr::summarise(text = paste(text, collapse = " "))

aggregate115 <- quanteda::corpus(aggregate115,
                                 docid_field = "year_month_chamber")
```

## Challenge 2

**Easy mode** Explore the context of the words *gun* and *firearm* with the
`kwic()` function. Are there any issues with the default setting?
Treat them as `glob` to capture plural forms.

```{r answers2_easy}
guns <- quanteda::kwic(corpus115, "gun*", window = 5, valuetype = "glob")
head(guns, 50)

firearms <- quanteda::kwic(corpus115, "firearm*", window = 5, valuetype = "glob")
head(firearms, 50)
```

**Medium** Now plot the kwic objects on the aggregated corpus ot explore when and
where these issues were discussed in the US Congress in the last two years.

```{r plot1}
quanteda::textplot_xray(kwic(aggregate115, "gun*", window = 5, valuetype = "glob"))
```

**Advanced** To make the previous plot a bit nicer, order the aggregated texts
by year, month, chamber. To do this, you would need to modify the ordering of
levels in the `aggregate115` dataset.

```{r answer2_advanced}
# Let us start from automatically creating the levels of text name
years <- rep(rep(c("2017", "2018"), times = c(12,11)), times = 2)
months <- rep(as.character(seq(1,12)), times = 2)
# Remove the last element as we don't have data for December 2018
months <- rep(months[-length(months)], times = 2)
chamber <- rep(c("H", "S"), each = 23)
# As Map() returns list we need to flatten it into vector
lev <- unlist(Map(function(x, y, z) paste(x, y, z, sep = "-"), years, months, chamber))

aggregate115 <- congress115 %>%
  dplyr::mutate(year = as.character(lubridate::year(date)),
                month = as.character(lubridate::month(date)),
                year_month_chamber = paste(year, month, chamber, sep = "-")) %>%
  dplyr::mutate(year_month_chamber = factor(year_month_chamber,
                                            levels = lev)) %>%
  dplyr::group_by(year_month_chamber) %>%
  dplyr::summarise(text = paste(text, collapse = " "))

aggregate115 <- quanteda::corpus(aggregate115,
                                 docid_field = "year_month_chamber")
```

```{r plot2}
quanteda::textplot_xray(kwic(aggregate115, "gun*", window = 5, valuetype = "glob"))
```

The gun debate, seemingly, intensified around February-March 2018 in both chambers,
in the aftermath of the number of school shootings. While Las Vegas shooting
in October 2017 saw a spike in frequency of references to guns in Congress, it
was less prominent than the one in early 2018.

**Subject Expert** Create a dictionary related to gun violence and firearms
control and apply to Congress debates.
